---
title: "Lesson8"
author: "Emi"
date: "2026-06-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load libraries
```{r}
library(tidyverse)
library(skimr)
```

#Load data
```{r}
chum <- read_csv("~/R Studio/r-collab/ClassNotes/Lesson7/20250730_AYKSSI_ASL_chum_export.csv")

skim(chum)

```

#Warm up:

Subset dataset to only include collections after 2010 located above 64 degrees latitude. Pipe this to ggplot to visualize results.
```{r}

chum %>%
  filter(CollectionYear > 2010, StartLatitudeDD > 64) %>%
  select(CollectionYear, StartLatitudeDD)

```

#Mutate()

```{r}
chum_asl <- chum %>%
  filter(is.na(AgeYRS) == F)
# estimate growth rate

chum_asl %>% mutate(growthrate = LengthMM/ AgeYRS)
```
```{r}
chum_asl %>% mutate(growthrate = LengthMM/ AgeYRS, .after = LengthMM)
```

#arrange()

We can sort our dataset by the 10 highest conc. Default organizes by 10 lowest first
```{r}

chum %>% 
  arrange(DNAngul) %>%
  select(ABLG, ExtractionPlate, ExtractionWell, DNAngul)
#if we want the highest conc.
chum %>% arrange(desc(DNAngul)) %>% select(ABLG, ExtractionPlate, ExtractionWell, DNAngul)

```

How many indvs are longer than 650mm and which pop did the largest indv come from
```{r}

```


#Summarize() and group_by()

Obtain mean conc. by pop. (can use na.rm for the mean_conc() operation for removing NA values in summarize())
```{r}

```



