---
title: "Lesson8"
author: "Emi"
date: "2026-06-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load libraries
```{r}
library(tidyverse)
library(skimr)
```

##Load data
```{r}
chum <- read_csv("~/R Studio/r-collab/ClassNotes/Lesson7/20250730_AYKSSI_ASL_chum_export.csv")

skim(chum)

```

##Warm up:

Subset dataset to only include collections after 2010 located above 64 degrees latitude. Pipe this to ggplot to visualize results.
```{r}

chum %>%
  filter(CollectionYear > 2010, StartLatitudeDD > 64) %>%
  select(CollectionYear, StartLatitudeDD)

```

##Mutate()

```{r}
chum_asl <- chum %>%
  filter(is.na(AgeYRS) == F)
# estimate growth rate

chum_asl %>% mutate(growthrate = LengthMM/ AgeYRS)
```
```{r}
chum_asl %>% mutate(growthrate = LengthMM/ AgeYRS, .after = LengthMM)
```

Excercise 2

Get the growthrate of males (1) and females (2) separately in two new columns

```{r}
chum %>% mutate(growthrate = LengthMM/ AgeYRS, .after = LengthMM) %>% 
  mutate(FemaleGrowthrate = growthrate, Sex = 2, .after = growthrate) %>% 
  mutate(MaleGrowthrate = case_when(growthrate, Sex = 1, .after = FemaleGrowthrate) %>% 
  filter(is.na(FemaleGrowthrate) == F) %>% 
   filter(is.na(MaleGrowthrate) == F)
```




##arrange()

We can sort our dataset by the 10 highest conc. Default organizes by 10 lowest first
```{r}

chum %>% 
  arrange(DNAngul) %>%
  select(ABLG, ExtractionPlate, ExtractionWell, DNAngul)
#if we want the highest conc.
chum %>% arrange(desc(DNAngul)) %>% select(ABLG, ExtractionPlate, ExtractionWell, DNAngul)

```

How many indvs are longer than 650mm and which pop did the largest indv come from
```{r}
View(chum)
skim(chum)
chum %>% filter(LengthMM > 650) %>% select(LengthMM, FreshwaterWatershed)

chum %>% filter(LengthMM > 650) %>% arrange(desc(LengthMM))

#there are 5 individuals longer than 650mm, but the longest indv came from Henshaw Creek-early at 670mm.
```


##Summarize() and group_by()

Obtain mean conc. by pop. (can use na.rm for the mean_conc() operation for removing NA values in summarize())
```{r}
#one way to find the mean of all indvs
chum %>% filter(is.na(DNAngul) == F) %>% summarize(mean_conc = mean(DNAngul))

#a different way to find the mean:
mean(chum$DNAngul, na.rm = T)

chum %>% 
  filter(is.na(DNAngul) == F) %>% 
  group_by(FreshwaterWatershed) %>% 
  summarize(mean_conc = mean(DNAngul))
#30 pops generate out of 53 because not all pops have DNA conc data

chum_dat %>% 
  #na.rm =T serves the same purpose as "filter(is.na(DNAngul) == F)" %>% 
  #here is another way to find the mean conc per pop, without filter
  group_by(FreshwaterWatershed) %>% 
  summarize(mean_conc = mean(DNAngul, na.rm = T))
```

##From Carl's Quarto document:

#### exercise

What is the largest fish caught at the earliest year?

```{r}
chum %>% arrange(desc(LengthMM)) %>% arrange(CollectionYear) %>% 
  filter(is.na(LengthMM) == F) %>% 
  select(ABLG, LengthMM, CollectionYear)

#the largest fish caught at the earliest year is ABLG #49272	@670mm in 2021 (the earliest year with length #value input)
```


## summarize() and group_by()

We can use summarize to calculate summary statistics for the whole dataset.

We can pair it with group_by() to get more useful statistics

mean DNA concentration per population:

```{r}
chum_dat %>% filter(is.na(DNAngul) == F) %>% summarise(mean_conc = mean(DNAngul))

mean(chum_dat$DNAngul, na.rm = T)

chum_dat %>% 
  filter(is.na(DNAngul) == F) %>% 
  group_by(FreshwaterWatershed) %>% 
  summarise(mean_conc = mean(DNAngul))
# notice only 30 of 53 pops are represented here. Some pops dont have any DNA conc data
chum_dat %>% 
  #filter(is.na(DNAngul) == F) %>% 
  group_by(FreshwaterWatershed) %>% 
  summarize(mean_conc = mean(DNAngul, na.rm = T))
```

#### exercise

Which population has the largest mean length?

##### bonus exercise

Show the lengths of each population using ggplot2 as a boxplot.
