---
title: "assignment_3"
author: "Emi"
date: "2026-06-23"
output: html_document
---

# **Exercise 1. Corruption and human development**

---  

This exercise explores a dataset containing the human development index (`HDI`) and corruption perception index (`CPI`) of 173 countries across 6 different regions around the world: Americas, Asia Pacific, Eastern Europe and Central Asia (`East EU Cemt`), Western Europe (`EU W. Europe`), Middle East and North Africa and Noth Africa (`MENA`), and Sub-Saharan Africa (`SSA`). (Note: the larger `CPI` is, the less corrupted the country is perceived to be.)

<br>

First, we load the data using the following code.
<br>

```{r}
## Write your code here
```

<br><br>

**1.1 Show the first few rows of `economist_data`.**
<br>

```{r}
## Write your code here
```

<br><br>

**1.2 Expore the relationship between human development index (`HDI`) and corruption perception index (`CPI`) with a scatter plot as the following.**
<br>

```{r}
## Write your code here
```

<br><br>

**1.3 Make the color of all points in the previous plot red**
<br>

```{r}
## Write your code here
```

<br><br>

**1.4 Color the points in the previous plot according to the `Region` variable, and set the size of points to 2.**
<br>

```{r}
## Write your code here
```

<br><br>

**1.5 Set the size of the points proportional to `HDI.Rank`**
<br>

```{r}
## Write your code here
```

<br><br>

**1.6 Fit a smoothing line to all the data points in the scatter plot from Excercise 1.4**
<br>

```{r}
## Write your code here
```

<br><br>

**1.7 Fit a separate straight line for each region instead, and turn off the confidence interval.**
<br>

```{r}
## Write your code here
```

<br><br>

**1.8 Building on top of the previous plot, show each `Region` in a different facet.**
<br>

```{r}
## Write your code here
```

<br><br>

**1.9 Show the distribution of `HDI` in each region using density plot. Set the transparency to 0.5**
<br>

```{r}
## Write your code here
```

<br><br>

**1.10 Show the distribution of `HDI` in each region using histogram and facetting.**
<br>

```{r}
## Write your code here
```

<br><br>

**1.11 Show the distribution of `HDI` in each region using a box plot. Set the transparency of these boxes to 0.5. Also show data points for each country in the same plot. (Hint: `geom_jitter()` or `position_jitter()` might be useful.)**
<br>

```{r}
## Write your code here
```

<br><br>

**1.12 Show the count of countries in each region using a bar plot.**
<br>

```{r}
## Write your code here
```

<br><br>

**1.13 You have now created a variety of different plots of the same dataset. Which of your plots do you think are the most informative? Describe briefly the major trends that you see in the data.**
<br>

```{r}
## Write your code here
```

<br><br><br>

# **Exercise 2. Unemployment in the US 1967-2015**

---  


This excercise uses the dataset `economics` from the ggplot2 package. It was produced from US economic time series data available from http://research.stlouisfed.org/fred2. It descibes the number of unemployed persons (`unemploy`), among other variables, in the US from 1967 to 2015.
<br>

```{r}
## Write your code here
```

<br><br>

**2.1 Plot the trend in number of unemployed persons (`unemploy`) though time using the economics dataset shown above. And for this question only, hide your code and only show the plot.**
<br>

```{r}
## Write your code here
```

<br><br>

**2.2 Edit the plot title and axis labels of the previous plot appropriately. Make y axis start from 0. Change the background theme to what is shown below. (Hint: search for help online if needed)**
<br>

```{r}
## Write your code here
```

<br><br>
