---
title: "Lesson12"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(nycflights13)  # install.packages("nycflights13")
```


# Understanding nycflights13:
  +  nycflights13 contains four tibbles that are related to the flights table:
    - airlines
    - airports
    - planes
    - weather
  +For nycflights13:

    - flights connects to planes via a single variable, tailnum.

    - flights connects to airlines through the carrier variable.

    - flights connects to airports in two ways: via the origin and dest variables.

    - flights connects to weather via origin (the location), and year, month, day and hour (the time).

# 13.2.1 Exercises

### 1) Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

  + You would need the lat and long from the airports table along with the origin and destination from the flights table. Combining the airport and flights tables would help, but faa needs to be mutated to origin.
  
```{r}
flights_ex1 <- flights %>% 
  select(flight, origin, dest) %>% 
  inner_join(airports, c("origin" = "faa"))
```


### 2) I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

  + Airports has an faa column which is synonymous with the origin column in the weather table. That is how they are connected. To demonstrate their relationship in the diagram, there should be a long arrow connecting airports' faa with weather's origin.

### 3) weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

  + It would provide the destination weather conditions, as well, for each flight/location.

### 4) We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

  + The data would need to be represented with columns indicating the flight number, date, time, number of passengers, and percent capacity. The primary keys of that table would be year, month, day, time, and flight number to create a unique identity.
  
# Notes on keys:

# There are two types of keys:

  + A primary key uniquely identifies an observation in its own table. For example, planes$tailnum is a primary key because it uniquely identifies each plane in the planes table.

  + A foreign key uniquely identifies an observation in another table. For example, flights$tailnum is a foreign key because it appears in the flights table where it matches each flight to a unique plane.

  + A variable can be both a primary key and a foreign key. For example, origin is part of the weather primary key, and is also a foreign key for the airports table.

  + Once you’ve identified the primary keys in your tables, it’s good practice to verify that they do indeed uniquely identify each observation. One way to do that is to count() the primary keys and look for entries where n is greater than one:

```{r}
planes %>% 
  count(tailnum) %>% 
  filter(n > 1)

#> # ℹ 2 variables: tailnum <chr>, n <int>
```



```{r}
weather %>% 
  count(year, month, day, hour, origin) %>% 
  filter(n > 1)


#> # A tibble: 3 × 6
#>    year month   day  hour origin     n
#>   <int> <int> <int> <int> <chr>  <int>
#> 1  2013    11     3     1 EWR        2
#> 2  2013    11     3     1 JFK        2
#> 3  2013    11     3     1 LGA        2
```

# Practice binding rows from multiple lotr datasets:
```{r}
fship <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Fellowship_Of_The_Ring.csv")

ttow <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Two_Towers.csv")

rking <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Return_Of_The_King.csv")


# And try binding these separate dataframes together: two ways to do it

#1
lotr_untidy <- dplyr::bind_rows(fship, ttow, rking)
# View the 9 generated rows
head(lotr_untidy, 9)

#2
lotr_untidy <- base::rbind(fship, ttow, rking)
head(lotr_untidy, 9)

View(airports)
```
# Join functions
### To practice the join functions, we’ll subset the flights dataframe as follows

```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
flights2
```
```{r}
View(flights)
View(airlines)
flights <- flights

avg_delay_dest <- flights %>% select(dep_delay, arr_delay, dest) %>% full_join(dep_delay, arr_delay, by = "dest")
```

### 1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:
```{r}
library(maps) #install.packages("maps")

airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()

View(planes)

```

```{r, eval = FALSE}
# first subset the dataframe for easier use and call the object flight_delays
# sum the arrival and departure delays, create a new column for those with mutate. Then take the avg/mean of total delay by destination.

flight_delays <- flights %>% 
  select(dep_delay, arr_delay, dest, origin) %>% 
  mutate(total_delay = dep_delay + arr_delay) %>% 
  group_by(dest) %>% 
  summarize(total_delay, .by = )

mean(flight_delays$total_delay, na.rm = TRUE)
```

# try again
```{r}

# each depature and arrival delay represents a different location, so their means have to be calculated separately 

mean_flight_delays <- flight_delays %>% group_by(dest) %>% 
  summarise(mean_arr_delay = mean(arr_delay, na.rm = TRUE), mean_dep_delay = mean(dep_delay, na.rm = TRUE)) 

mean_flight_delays %>% inner_join(airports, c("dest" = "faa")) %>% 
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(aes(color = mean_dep_delay), size = 3) +
    scale_color_viridis_c(option = "C + B") +
    coord_quickmap()
``` 

