---
title: "Lesson9"
author: "Emi"
date: "2026-07-07"
output: html_document
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))

coronavirus <- read_csv('https://raw.githubusercontent.com/RamiKrispin/coronavirus/master/csv/coronavirus.csv', col_types = cols(province = col_character()))

# Has it been updated? Check the latest date?
max(coronavirus$date)

View(coronavirus)
```

###Let’s start with summarizing the total number of cases by type as of the most recent day in the dataset

```{r}
names(coronavirus)

cases <- coronavirus %>% group_by(type) %>% 
  summarize(cases = sum(cases))
kable(cases) # kable() just provides a nice output for the table

cases
View(cases)

```


###Now, let’s plot the history of the daily counts of new confirmed cases worldwide
```{r}
# We first have to summarize the data, then plot those summary statistics
# Note that we can pipe dplyr output directly into ggplot
coronavirus %>%  
  filter(type == "confirmed") %>% 
  group_by(date) %>% 
  summarize(total_cases = sum(cases)) %>% 
  ggplot(mapping = aes(x = date, y = total_cases)) +
  geom_line()

# If we want to play around with different geoms, we can store dplyr data processing steps and initiation of the ggplot as object gg_base so that we don’t need to retype it each time

# define a base map dataset
gg_base <- coronavirus %>%  
  filter(type == "confirmed") %>% 
  group_by(date) %>% 
  summarize(total_cases = sum(cases)) %>% 
  ggplot(mapping = aes(x = date, y = total_cases))

# line
gg_base +
  geom_line()

# point
gg_base +
  geom_point()

# line and point
gg_base +
  geom_line() + 
  geom_point()

# area
gg_base +
  geom_area()

```

```{r}
gg_base +
  geom_col(color="red")
```
### We can change the aesthetics of elements in a ggplot graph by adding arguments within the layer where that element is created. Some common arguments we’ll use first are:

####color =: update point or line colors
####fill =: update fill color for objects with areas
####linetype =: update the line type (dashed, long dash, etc.)
####pch =: update the point style
####size =: update the element size (e.g. of points or line thickness)
####alpha =: update element opacity (1 = opaque, 0 = transparent)
#####Building on our first line graph, let’s update the line color to “purple” and make the line type “dashed”:

```{r}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```
###How do we know which color names ggplot will recognize? If you google “R colors ggplot2” you’ll find a lot of good resources. Here’s one: SAPE ggplot2 colors quick reference guide

###Now let’s update the point, style and size of points on our previous scatterplot graph using color =, size =, and pch = (see ?pch for the different point styles, which can be further customized).
```{r}
gg_base + 
  geom_point(color = "purple",
             pch = 17,
             size = 4,
             alpha = 0.5)
```

###Example: Create a ggplot scatterplot graph where the size and color of the points change based on the number of cases, and make all points the same level of opacity (alpha = 0.5). Notice the aes() around the size = and color = arguments.

###Note: this is just for illustration of the functionality only - we are showing the same information in redundant ways, which is typically not helpful or necessary. Avoid excessive / overcomplicated aesthetic mapping in data visualization.

```{r}
gg_base + 
  geom_point(
    aes(size = total_cases,
        color = total_cases),
    alpha = 0.5
  )

#In the example above, notice that the two arguments that do depend on variables are within aes(), but since alpha = 0.5 doesn’t depend on a variable then it is outside the aes() but still within the geom_point() layer.
```

###While every element of a ggplot graph is manually customizable, there are also built-in themes (theme_*()) that you can add to your ggplot code to make some major headway before making smaller tweaks manually.

###We talked about this briefly a few classes ago, but let’s explore a little further. Here are a few to try today (but also notice all the options that appear as we start typing theme_ into our ggplot graph code!):

####theme_light()
####theme_minimal()
####theme_bw()
####Also, check out more examples by scrolling down here. Pick one that you like and update your previous plot.

#####Here, let’s update our previous graph with theme_minimal():
```{r}
gg_base + 
  geom_point(
    aes(size = total_cases,
        color = total_cases),
    alpha = 0.5
  ) +
  theme_minimal()
```
##Updating axis labels and titles
###Use labs() to update axis labels, and add a title and/or subtitle to your ggplot graph.
```{r}
gg_base +
  geom_line(linetype = "dotted") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Total confirmed cases",
    title = "Daily counts of new Coronavirus cases recorded",
    subtitle = "Global sums"
  )
```
#Now, let’s split the case counts out by country


###Your turn
####Take a minute to think about how you would generate a plot with a separate line showing the daily reports of new confirmed cases in each country.

```{r}
# We need columns date, country, and type = confirmed

coronavirus %>% 
```






