---
title: "Lesson10"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
install.packages("janitor", repos = "https://cloud.r-project.org")
##in the future, install the package in the console or else the file won't knit
# or you will have to indicate a cloud mirror like I have above (repos)
```
```{r}
# Attach packages
library(tidyverse)
library(janitor)  ## install.packages("janitor")
```

Import datasets
```{r}
lotr <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/lotr_tidy.csv")
```

```{r}
fship <- read.csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Fellowship_Of_The_Ring.csv")


```

```{r}
ttow <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Two_Towers.csv")
```

```{r}
rking <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Return_Of_The_King.csv")
```
#gather these data
```{r}

lotr_untidy <- bind_rows(fship, ttow, rking)
str(lotr_untidy)

```
#tidy up the data by combining the gender columns
```{r}
lotr_tidy <-
  pivot_longer(lotr_untidy, c(Male, Female), names_to = 'Gender', values_to = 'Words')

lotr_tidy
```

```{r}
#4. Write the tidy data to a delimited file
#Now we write this multi-film, tidy dataset to file for use in various downstream scripts for further analysis and visualization.

#write_csv(lotr_tidy, path = "../datasets/lotr_tidy.csv")
#getting one variable per race
```




```{r}
lotr_tidy %>%  
  pivot_wider(names_from = Race, values_from = Words)
```
#Separating gender columns
```{r}
lotr_tidy %>% 
  pivot_wider(names_from = Gender, values_from = Words)
```
#Unite gender and race columns
```{r}
# let's get one variable per combo of Race and Gender
lotr_tidy %>% 
  unite(Race_Gender, Race, Gender) %>% 
  pivot_wider(names_from = Race_Gender, values_from = Words)
```

##############################################################################

CORONAVIRUS

Applying this to our Coronavirus dataset
####Let’s now return to our Coronavirus dataset. Let’s remind ourselves of it’s structure

```{r}
coronavirus <- read_csv('https://raw.githubusercontent.com/RamiKrispin/coronavirus/master/csv/coronavirus.csv', col_types = cols(province = col_character()))

coronavirus
```

###Last class, we visualized the global case counts date

```{r}
coronavirus %>% 
  group_by(date, type) %>% filter(cases > 0) %>%
  summarize(cases = sum(cases)) %>%
  ggplot() +
  geom_col(aes(x = date, y = cases, fill = type))
```

###Let’s see how we would do that if the data had been in a wider format.
####Your turn
####Convert the coronavirus dataset to a wider format where the confirmed cases, deaths and recovered cases are shown in separate columns.
```{r}
corona_wide <- coronavirus %>% filter(cases > 0) %>%
  pivot_wider(names_from = type, values_from = cases)
```

####Now how do we reproduce the barchart of total cases per day broken down by type?

####And how would we plot the daily counts of different case types within a country? With the long format this is easy:

```{r}
coronavirus %>% filter(cases > 0) %>%
  filter(country == "US") %>% 
  ggplot() +
  geom_line(aes(x = date, y = cases, color = type))
```

####How would we do this with the coronavirus_wide format? That would be much more difficult

####As mentioned above, however, there are plot types where the wide format provides the best ####input. For example, in Slack, I showed the example of plotting the total death count per ####country against the total count of confirmed cases. It would be cumbersome to pull these out of ####the long format because in ggplot we are mapping variables to aesthetics and now we want to map ####different levels of a variable to different aesthetics. So let’s make those different levels ####separate variables by widening the data.

```{r}
coronavirus_ttd <- coronavirus %>% 
  group_by(country, type) %>%
  summarize(total_cases = sum(cases)) %>%
  pivot_wider(names_from = type, values_from = total_cases)

# Now we can plot this easily
ggplot(coronavirus_ttd) +
  geom_label(mapping = aes(x = confirmed, y = death, label = country))
```
###Using janitor::clean_names()

####We’ve noticed in the coronavirus dataset that the variable names have a very inconsistent format. It’s not a problem per se for our ability to work with this dataframe, but it’s a bit annoying to look at.


```{r}
colnames(coronavirus)
```
#If we wanted to then use these columns (which we probably would, since we created them), we could clean the names to get them into more coder-friendly lower_snake_case with janitor::clean_names():

```{r}
coronavirus <- coronavirus %>% 
  clean_names()

names(coronavirus)
```

###And there are other case options in clean_names(), like:

###### “snake” produces snake_case (the default)
##### “lower_camel” or “small_camel” produces lowerCamel
##### “upper_camel” or “big_camel” produces UpperCamel
##### “screaming_snake” or “all_caps” produces ALL_CAPS
##### “lower_upper” produces lowerUPPER
##### “upper_lower” produces UPPERlower