---
title: "Lesson12"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(nycflights13) # install.packages("nycflights13")
library(dplyr)
library(ggplot2) # is already in tidyverse but just a reminder
```


# Understanding nycflights13:
+  nycflights13 contains four tibbles that are related to the flights table:
- airlines
- airports
- planes
- weather
+For nycflights13:

- flights connects to planes via a single variable, tailnum.

- flights connects to airlines through the carrier variable.

- flights connects to airports in two ways: via the origin and dest variables.

- flights connects to weather via origin (the location), and year, month, day and hour (the time).

# 13.2.1 Exercises

### 1) Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

+ You would need the lat and long from the airports table along with the origin and destination from the flights table. Combining the airport and flights tables would help, but faa needs to be mutated to origin.

```{r}
flights_ex1 <- flights %>% 
  select(flight, origin, dest) %>% 
  inner_join(airports, c("origin" = "faa"))
```


### 2) I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

+ Airports has an faa column which is synonymous with the origin column in the weather table. That is how they are connected. To demonstrate their relationship in the diagram, there should be a long arrow connecting airports' faa with weather's origin.

### 3) weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

+ It would provide the destination weather conditions, as well, for each flight/location.

### 4) We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

+ The data would need to be represented with columns indicating the flight number, date, time, number of passengers, and percent capacity. The primary keys of that table would be year, month, day, time, and flight number to create a unique identity.

# Notes on keys:

# There are two types of keys:

+ A primary key uniquely identifies an observation in its own table. For example, planes$tailnum is a primary key because it uniquely identifies each plane in the planes table.

+ A foreign key uniquely identifies an observation in another table. For example, flights$tailnum is a foreign key because it appears in the flights table where it matches each flight to a unique plane.

+ A variable can be both a primary key and a foreign key. For example, origin is part of the weather primary key, and is also a foreign key for the airports table.

+ Once you’ve identified the primary keys in your tables, it’s good practice to verify that they do indeed uniquely identify each observation. One way to do that is to count() the primary keys and look for entries where n is greater than one:

```{r}
planes %>% 
  count(tailnum) %>% 
  filter(n > 1)

#> # ℹ 2 variables: tailnum <chr>, n <int>
```



```{r}
weather %>% 
  count(year, month, day, hour, origin) %>% 
  filter(n > 1)


#> # A tibble: 3 × 6
#>    year month   day  hour origin     n
#>   <int> <int> <int> <int> <chr>  <int>
#> 1  2013    11     3     1 EWR        2
#> 2  2013    11     3     1 JFK        2
#> 3  2013    11     3     1 LGA        2
```

# Practice binding rows from multiple lotr datasets:
```{r}
fship <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Fellowship_Of_The_Ring.csv")

ttow <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Two_Towers.csv")

rking <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Return_Of_The_King.csv")


# And try binding these separate dataframes together: two ways to do it

#1
lotr_untidy <- dplyr::bind_rows(fship, ttow, rking)
# View the 9 generated rows
head(lotr_untidy, 9)

#2
lotr_untidy <- base::rbind(fship, ttow, rking)
head(lotr_untidy, 9)

View(airports)
```
# Join functions
### To practice the join functions, we’ll subset the flights dataframe as follows

```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
flights2
```
```{r, eval=F}
View(flights)
View(airlines)
flights <- flights

avg_delay_dest <- flights %>% select(dep_delay, arr_delay, dest) %>% full_join(dep_delay, arr_delay, by = "dest")
```

### 1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:
```{r}
library(maps) #install.packages("maps")

airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
  borders("state") +
  geom_point() +
  coord_quickmap()

View(planes)

```

```{r, eval = FALSE}
# first subset the dataframe for easier use and call the object flight_delays
# sum the arrival and departure delays, create a new column for those with mutate. Then take the avg/mean of total delay by destination.

#flight_delays <- flights %>% 
  #select(dep_delay, arr_delay, dest, origin) 


 # mutate(total_delay = dep_delay + arr_delay) %>% 
  #group_by(dest) %>% 
  #summarize(total_delay, .by = )

#mean(flight_delays$total_delay, na.rm = TRUE)
```

# try again
```{r}

# each depature and arrival delay represents a different location, so their means have to be calculated separately 

flight_delays <- flights %>% 
  select(dep_delay, arr_delay, dest, origin, tailnum) %>% 
  mutate(total_delay = dep_delay + arr_delay)


mean_flight_delays <- flight_delays %>% group_by(dest) %>% 
  summarise(mean_arr_delay = mean(arr_delay, na.rm = TRUE), mean_dep_delay = mean(dep_delay, na.rm = TRUE)) 

mean_flight_delays %>% inner_join(airports, c("dest" = "faa")) %>% 
  ggplot(aes(lon, lat)) +
  borders("state") +
  geom_point(aes(color = mean_dep_delay), size = 3) +
  scale_color_viridis_c(option = "C + B") +
  coord_quickmap()
``` 



yoooo  idk how to make this work
```{r, eval=F}
library(usmap)
# map <- plot_usmap()

flight_delays <- flights %>% 
  select(dep_delay, arr_delay, dest, origin, tailnum) %>% 
  mutate(total_delay = dep_delay + arr_delay)

#mean_flight_delays %>% inner_join(airports, c("dest" = "faa")) %>% 
 # ggplot(aes(lon, lat)) +
 # borders("state") +
 # geom_point(aes(color = mean_dep_delay), size = 3) +
 # scale_color_viridis_c(option = "C")
```

```{r, eval= FALSE}

# my attempt
#mean_flight_delays %>% inner_join(airports, c("dest" = "faa")) %>% 
 # plot_usmap(regions = "states", data = mean_flight_delays, values = "dest") %>% 
 #  ggplot(aes(lon, lat)) +
 # borders("state") +
 # geom_point(aes(color = mean_dep_delay), size = 3) +
 # scale_color_viridis_c(option = "C")
```
this code doesn't really work oh well
```{r, eval = FALSE}
# 1. Join your data and filter out airports outside the US map boundaries
map_data <- mean_flight_delays %>% 
  inner_join(airports, by = c("dest" = "faa")) %>% 
  filter(!is.na(lon) & !is.na(lat))

# 2. Transform the raw longitude/latitude points into the usmap projection
transformed_data <- usmap_transform(map_data, input_cols = c("lon", "lat"))

# 3. Build the plot using the transformed coordinates
plot_usmap(regions = "states", color = "gray80", fill = "white") +
  geom_point(data = map_data, 
             aes(x = lon, y = lat, color = mean_dep_delay), 
             size = 3) +
  scale_color_viridis_c(option = "plasma") + # 'plasma' corresponds to option C
  labs(color = "Mean Dep Delay")
```

Example 2:

# Add the location of the origin and destination (i.e. the lat and lon) to flights.
```{r}
locations <- airports %>% select(lon, lat, faa)

flight_org_dest <- flights %>% select(origin, dest, flight, year, month, day, dep_time)

# dont even need to do above step, could just create flight_org from flights itself

flight_org <- flight_org_dest %>% select(origin, flight, year, month, day, dep_time)

flight_dest <- flight_org_dest %>% select(dest, flight, year, month, day, dep_time)

flight_org_loc <- flight_org %>% inner_join(locations, c("origin" = "faa")) %>%          
  rename(org_lon = lon, org_lat = lat)

flight_dest_loc <- flight_dest %>% inner_join(locations, c("dest" = "faa")) %>%          
  rename(dest_lon = lon, dest_lat = lat)

flight_locations <- flight_org_loc %>% full_join(flight_dest_loc, 
                                                 by = join_by(flight, year, month, day, dep_time),
                                                 relationship = "many-to-many")


# there could be a much simpler way of doing this, potentially using full_join and
# left_join without making a ton of new objects. IDK we are learning here.
```

# THE LAST EXAMPLE PROBLEM:
### Is there a relationship between the age of a plane and its delays?

```{r}
View(planes)

library(ggplot2)

flight_delays <- flights %>% 
  select(dep_delay, arr_delay, dest, origin, tailnum) %>% 
  mutate(total_delay = dep_delay + arr_delay)

flight_delays[c("dep_delay", "arr_delay")][is.na(flight_delays[c("dep_delay", "arr_delay")])] <- 0

clean_planes <- planes %>% select(tailnum, year) %>% rename(Manufactured_year = year) %>% 
  arrange(Manufactured_year, decreasing = FALSE) %>% na.omit

rel_plane_delay <- flight_delays %>% inner_join(clean_planes, by = "tailnum") %>% 
  arrange(Manufactured_year, decreasing = FALSE)

# seems to work but I'd love to make better visuals to display the relationship between age of plane and delay. ggplot is not my friend today.
ggplot(rel_plane_delay) +
  geom_point(aes(x = Manufactured_year, y = total_delay))
  
  
# IDK how to add geom_smooth() to make it right, this is so hard

# weird lines ahh what is this even
ggplot(rel_plane_delay) +
  geom_line(aes(x = Manufactured_year, y = total_delay))

## ahh scary this is awful
ggplot(rel_plane_delay) +
  geom_boxplot(mapping = aes(x = Manufactured_year, y = total_delay, group=total_delay ))


  

```

### Answer: yes and no, it depends. We don't have enough information on what causes the delays. It could be a result of the plane and its age, but it could also be a plethora of other factors. The point plot shows more modern planes are used more often, which makes sense. But reasoning data is not available. All planes have lots of delays. I would like to plot the data differently and do some statistical tests, but in this current moment I do not know how and would like to move on.
